### 2016-10-27 capability-machine.tex
typos
Added suggestion for world and part of value relation

%Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{hs}}
\newcommand{\hv}{\var{hv}}
\newcommand{\rv}{\var{rv}}
\newcommand{\lv}{\var{lv}}
\newcommand{\gl}{g}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{hs}}
\newcommand{\heap}{\var{heap}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}

@@

@@

@@

%View
\newcommand{\perma}{\mathrm{permanent}}
\newcommand{\temp}{\mathrm{temporary}}

@@

@@

\subsubsection{Continuation Relation}
\label{subsubsec:continuation-relation}
The continuation relation is used in the definition of the expression relation. The continuation relation ensures that if you continue execution through a continuation, then it will result in a good result according to the world.
\begin{align*}
  \stdkr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
                     \begin{aligned}[t]
                       & \npair{c} \in \stdvr(W) \land \\
                       &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{reg} \in \stdrr(\stdvr)(W') \ldotp \\
                       & \quad \npair[n']{(reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
                 \end{aligned}
\end{align*}
We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.

@@

\subsubsection{Value Relation}
\label{subsubsec:value-relation}
The value relation, is defined as follows:
\begin{align*}
  \stdvr & :  (\Worlds \monnefun \UPred{Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
              \begin{aligned}[t]
  & \{ \npair{i} \mid i \in \ints \} 
    \union \\
         & \{ \npair{\stdcap[\noperm] }  \} 
           \union \\
         & \{ \npair{\stdcap[\readonly] } \mid n \in \readCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\readwrite] } \mid n \in \writeCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\exec]} \mid \\
         & \quad n \in \readCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \}
           \union \\
           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
         & \{ \npair{\stdcap[\entry]} \mid n \in \entryCond{}(V)(\start,\addrend,\addr,W) \} \union \\
         & \{ \npair{\stdcap[\rwx]} \mid \\
         & \quad n \in \writeCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\rwx,W) \}
       \end{aligned}
\end{align*}

@@

\subsection{Logical Relation}
\lau{To be deleted when incorporated in the rest of the text.}
\subsubsection{Worlds}
\lau{disclaimer: the following is my current perception of what is going on}
The worlds here are almost the same as the ones we defined in the case without local capabilities in Section~\ref{subsec:worlds}. The difference is that here our region keeps track of a view, namely whether it belongs to the \emph{permanent} or \emph{temporary} view of the current memory. The permanent part signifies what can be kept after a return to a caller whereas the temporary part will be lost after returning to a caller.

Define $\Views$ to be
\[
  \Views \defeq \{\perma, \temp \}
\]

Assume a preordered \cofe{} $\Wor$ and an isomorphism $\xi$ such that:
\[
  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
\]

And the worlds are
\[
  \Worlds \defeq \nats \finparfun \Regions
\]

>>>>

%Variables
\newcommand{\var}[1]{\mathit{#1}}
\newcommand{\hs}{\var{hs}}
\newcommand{\hv}{\var{hv}}
\newcommand{\rv}{\var{rv}}
\newcommand{\lv}{\var{lv}}
\newcommand{\gl}{\var{g}}
\newcommand{\pc}{\mathit{pc}}
\newcommand{\pcreg}{\mathrm{pc}}
\newcommand{\addr}{\var{a}}
\newcommand{\offset}{\var{offset}}
\newcommand{\word}{\var{w}}
\newcommand{\start}{\var{base}}
\newcommand{\addrend}{\var{end}}
\newcommand{\pwlv}{\var{pwl}}
\newcommand{\mem}{\var{mem}}
\newcommand{\reg}{\var{reg}}
\newcommand{\heapseg}{\var{hs}}
\newcommand{\heap}{\var{heap}}
\newcommand{\mode}{\var{mode}}
\newcommand{\perm}{\var{perm}}
\newcommand{\permp}{\var{permPair}}
\newcommand{\roll}{\var{roll}}
\newcommand{\instr}{\var{instr}}
\newcommand{\stdcap}[1][(\perm,\gl)]{\left(#1,\start,\addrend,\addr \right)}

@@

\newcommand{\revokeTemp}[1]{\plainfun{revokeTemp}{#1}}
\newcommand{\erase}[2]{\floor*{#1}_{#2}}

@@

%Views
\newcommand{\plainview}[1]{\mathrm{#1}}
\newcommand{\perma}{\plainview{perm}}
\newcommand{\temp}{\plainview{temp}}
\newcommand{\revoked}{\plainview{revoked}}

@@

@@

\section{Logical Relation'}
\subsection{Worlds}
Define the domain $\Views$ as follows
\[
  \Views = \{\perma,\temp,\revoked\}
\]
Using the ordering where $\revoked \future \temp$ and $\future$ is otherwise reflexive, and a family of equivalences where everything is equal at 0 and othervise the equivalences are ``normal'' equality. We claim this is a preordered \cofe{}. %TODO prove this claim! And show the new "region" is a preordered cofe with the claimed ordering.

Assume $\Wor$ is a preordered \cofe{} and $\xi$ is an isomorphism such that
\[
  \xi : \Wor \cong \blater (\nats \finparfun (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments}))))
\]
The regions are here
\[
  \Regions = (\Views \times \States \times \Rel \times (\States \fun (\Wor \monnefun \UPred{\HeapSegments})))
\]
And the worlds are
\[
  \Worlds = \RegionNames \finparfun \Regions
\]
The ordering of worlds is

 \begin{align*}
 W' \future W & & \text{ iff } & &
   \begin{gathered}
     \dom(W') \supseteq \dom(W)\\ 
     \text{ and } \\
     \forall r \in \dom(W) \ldotp W'(r) \future W(r)
   \end{gathered}
 \end{align*}
and the ordering of regions is
\begin{gather*}
  (v_2,s_2,\phi_2,H_2) \future (v_1,s_1,\phi_1,H_1) \\
  \text{ iff } \\ 
    \begin{gathered}
      (v_2,v_1) = (\revoked,\revoked) \text{ and } (s_1,\phi_1,H_1) = (s_2,\phi_2,H_2) \\
      \text{or} \\
      (v_2,v_1) \in \{(\revoked,\temp), (\temp,\temp), (\perma,\perma) \} \text{ and } \\
      (s_1,s_2) \in \phi_2 \text{ and } (\phi_1,H_1) = (\phi_2,H_2)
    \end{gathered} 
\end{gather*}
The idea here is that the region is ``frozen'' after it has been revoked.

Erase all but one view in a world:

\begin{align*}
  \erase{W}{v} \defeq \lambda r \ldotp 
  \begin{cases}
    W(r) & W(r).v = v\\
    \undefined & \text{otherwise}
  \end{cases}
\end{align*}

Standard region for when writing locally is permitted:
\begin{align*}
  &\iota_{\start,\addrend}^{\var{pwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{pwl}}) \\ 
  &H^{\var{pwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \npair[n-1]{\hs(\addr)} \in \stdvr(\xi(\hat{W}))
    \end{aligned}
  \right\}
\end{align*}

Rewoking all temporary regions:

\begin{align*}
  \revokeTemp{} & : \Worlds \fun \Worlds \\
  \revokeTemp{W} & \defeq \lambda r \ldotp 
                   \begin{cases}
                     (\revoked,s,\phi,H) & W(r) = (\temp,s,\phi,H) \\
                     W(r)                & \text{otherwise}
                   \end{cases}
\end{align*}

A standard region for when write local is not allowed:

\begin{align*}
  &\iota_{\start,\addrend}^{\var{nwl}} \defeq \lambda v \ldotp (v,(\start,\addrend),=,H^{\var{nwl}}) \\ 
  &H^{\var{nwl}} \; (\start,\addrend) \; \hat{W} \defeq \left\{\npair{\hs} \mid
    \begin{aligned}
      &\dom(\hs) = [\start,\addrend] \land \\
      &\forall \addr \in [\start,\addrend] \ldotp \\
      &\quad \npair[n-1]{\hs(\addr)} \in \stdvr(\revokeTemp{\xi(\hat{W})})
    \end{aligned}
  \right\}
\end{align*}

\begin{align*}
    & \writeCond{}(\iota,\var{regions})(\start,\addrend,W) =  \\
    & \quad \begin{aligned}[t]
              \{ n \mid & \;\exists r \in \var{regions} \ldotp \\
              & \;\quad \exists [\start',\addrend'] \supseteq [\start,\addrend] \ldotp \\
              & \;\qquad W(r)\nequal[n-1] \iota_{\start',\addrend'}(W(r).v) \}
            \end{aligned}
\end{align*}

Now define the value relation as follows:
\begin{align*}
  \stdvr &\defeq \lambda \; V \ldotp 
           \begin{aligned}[t]
             & \{ \npair{i} \mid i \in \ints \} 
             \union \\
             & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
             \union \\
             & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \} 
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{nwl}},\dom(W))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwrite,\glob)] } \mid  \\            
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\dom(\erase{W}{\perma}))(\start,\addrend,W) \}
             \union \\
             & \{ \npair{\stdcap[(\readwritel,\local)] } \mid \\
             & \quad n \in \writeCond{}(\iota^{\var{pwl}},\var{regions})(\start,\addrend,W) \\
             & \quad \text{for }\var{regions} = \dom(\erase{W}{\temp}) \union \dom(\erase{W}{\revoked})\}
             \union \\
             & \{ \npair{\stdcap[(\exec,\gl)]} \mid \}
             \union \\
             & \{ \npair{\stdcap[(\entry,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \} \union \\
             & \{ \npair{\stdcap[(\rwlx,\local)]} \mid \}
           \end{aligned}
\end{align*}

@@

\subsubsection{Continuation Relation}
\label{subsubsec:continuation-relation}
The continuation relation is used in the definition of the expression relation. The continuation relation ensures that if you continue execution through a continuation, then it will result in a good result according to the world.
\begin{align*}
  \stdkr \quad : & \quad (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdkr \defeq & \lambda \stdvr \ldotp \lambda W \ldotp\{ \npair{c} \mid
                     \begin{aligned}[t]
                       & \npair{c} \in \stdvr(W) \land \\
                       &  \forall W' \future W, n' < n\ldotp %on the whiteboard, we had public future world.
                       \forall \heapSat[\hs]{n'}{W'} \ldotp \forall \reg, \npair[n']{\reg} \in \stdrr(\stdvr)(W') \ldotp \\
                       & \quad \npair[n']{(\reg\update{\pcreg}{\updatePcPerm{\var{c}}},\hs)} \in \observations(W') \}
                 \end{aligned}
\end{align*}
We require the continuation to be in the value relation, because it will have to be in the registers when it is invoked, so to argue that the register file is in the register-file relation, we need to know that the continuation is in the value relation.

@@

\subsubsection{Value Relation}
\label{subsubsec:value-relation}
The value relation, is defined as follows:
\begin{align*}
  \stdvr & :  (\Worlds \monnefun \UPred{\Words}) \nefun \Worlds \monnefun \UPred{\Words} \\
  \stdvr &\defeq \lambda \; V \ldotp \lambda \; W \ldotp 
              \begin{aligned}[t]
  & \{ \npair{i} \mid i \in \ints \} 
    \union \\
         & \{ \npair{\stdcap[\noperm] }  \} 
           \union \\
         & \{ \npair{\stdcap[\readonly] } \mid n \in \readCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\readwrite] } \mid n \in \writeCond{}(V)(\start,\addrend,W) \} \union \\
         & \{ \npair{\stdcap[\exec]} \mid \\
         & \quad n \in \readCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \}
           \union \\
           % For the entry case a capability is acceptable, if we can change it to an execute permission and put it as the pc and we otherwise have valid words in the registers, then the register file should be valid.
         & \{ \npair{\stdcap[\entry]} \mid n \in \entryCond{}(V)(\start,\addrend,\addr,W) \} \union \\
         & \{ \npair{\stdcap[\rwx]} \mid \\
         & \quad n \in \writeCond{}(V)(\start,\addrend,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\exec,W) \land \\
         & \quad n \in \execCond{}(V)(\start,\addrend,\rwx,W) \}
       \end{aligned}
\end{align*}

@@
